Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-11-19 paper/ifo-introduction.tex
small edits

In 2015, the first detection of gravitational waves from the merger of two black holes~\cite{GW150914} marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
% What are gws and gw detectors 
Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}, 
They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
The effect of gravitational waves is a change in lengths: a `stretching and squashing' of the distance between objects. 
Ground-based gravitational wave observatories such as the Advanced Laser Interferometer Gravitational wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance. 
Gravitational wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.

% CW 
To date, the network of gravitational wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}. 
However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
Rotating neutron stars are prime candidates for continuous wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
%where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
% we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
Scorpius X-1 is a prime LMXB target for continuous wave searches. 
%Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.

% changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
In this paper, we use a table-top Michelson interferometer as a toy gravitational wave detector designed to detect sound instead of gravitational waves. 
We then extend its use to an ``optical microphone'' (using light to capture sound) and present a range of example analysis techniques for educators to use. 
% re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
%% I don't like the use of ``freedom to explore''. 
% removed Additionally - I don't think it is neccessary - hannah
It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}. 
% This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
% Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational wave research groups around the world. 
This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}. 
These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational wave science to non-specialist audiences.
%\han{look up Sarah Farmer's project}

This paper is laid out as follows. 
In Section~\ref{sec:ifo}, we detail the table-top interferometer design. 
In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
% removed as there's not space to really say what this means here: to varying degrees of success.
This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples which can be used in the undergraduate laboratory. 
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
We collect the software and scripts used to produce this work in Appendix~\ref{app:code}.

>>>>

In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}, 
They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
The effect of gravitational waves is a change in lengths: a `stretching and squashing' of the distance between objects. 
Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance. 
These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.

% CW 
To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}. 
However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
Rotating neutron stars are prime candidates for continuous-wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
%where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
% we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
Scorpius X-1 is a prime LMXB target for continuous wave searches. 
%Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.

% changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
We then extend its use to an ``optical microphone'' (using light to capture sound) and present a range of example analysis techniques for educators to use. 
% re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
%% I don't like the use of ``freedom to explore''. 
% removed Additionally - I don't think it is neccessary - hannah
It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}. 
% This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
% Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational wave research groups around the world. 
This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}. 
These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.
%\han{look up Sarah Farmer's project}

This paper is laid out as follows. 
In Section~\ref{sec:ifo}, we detail the table-top interferometer design. 
In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
% removed as there's not space to really say what this means here: to varying degrees of success.
This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples which can be used in the undergraduate laboratory. 
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
We collect the software and scripts used to produce this work in Appendix~\ref{app:code}.

### 2020-11-19 paper/ifo-singleTone.tex
full grammar check before submission
small edits

Continuous-wave searches look for nearly monochromatic signals. In this section we consider a simple sinusoidal tone at a single, constant frequency. %should be single constant frequency *signal*?

@@

%The pixel intensity in the green-channel of the video (the webcam records in three colour channels: red, blue, and green) is used to approximate the total intensity of the interference pattern. 
%This is because the laser emits monochromatic green light at $532\,{\rm nm}$.
A tone with a frequency of $2.09\,{\rm Hz}$ is played through the speaker for one minute. 
The amplitude of the Fourier transform of the recovered signal is shown in Fig.~\ref{fig:webcam_spectrum}. 
% removed 'strong' as what does strong mean here?
We measured a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$. 
%Although this range doesn't contain our injected frequency, the error is not substantial and cannot be heard by the (untrained) human ear.
Two harmonics can also be seen at integer multiples of the peak frequency. 
The peaks at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ have amplitudes of $0.103$ and $0.098$ as a fraction of the height of the main peak, respectively. They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo} and Appendix~\ref{app:intensity_derivation}.

In the following section, we use a hidden Markov model technique, commonly used in continuous wave searches, to recover a slowly wandering frequency signal.

>>>>

Continuous-wave searches look for nearly monochromatic signals. In this section, we consider a simple sinusoidal tone at a single, constant frequency. %should be single constant frequency *signal*?

@@

A tone with a frequency of $2.09\,{\rm Hz}$ is played through the speaker for one minute. 
The amplitude of the Fourier transform of the recovered signal is shown in Fig.~\ref{fig:webcam_spectrum}. 
We measured a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$. 
Two harmonics can also be seen at integer multiples of the peak frequency. 
The peaks at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ have amplitudes of $0.103$ and $0.098$ as a fraction of the height of the main peak, respectively. They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo} and Appendix~\ref{app:intensity_derivation}.

In the following section, we use a hidden Markov model technique, commonly used in continuous-wave searches, to recover a slowly wandering frequency signal.

### 2020-11-19 paper/ifo-tableTopGWs.tex
small edits

% rephrased to mention the laser as the first component and shifted the arm lengths of LIGO and Virgo to later in the para when we discuss them. 
Gravitational wave detectors such as LIGO and Virgo are large, complex experiments. 
However, their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
In a Michelson interferometer, laser light is split by a beamsplitter into two perpendicular arms (see Fig.~\ref{fig:ifo_schematic_webcam}). 
Mirrors at the end of each arm reflect the two beams back to the beamsplitter where they recombine to produce an interference pattern, 
The resulting interference pattern is dependent on the relative distance travelled by the beams. 
Current generation gravitational wave observatories have kilometre-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
They can measure minuscule changes in distance due to gravitational waves, e.g. the first detection of a binary black hole merger produced a strain of $10^{-21}$, corresponding to a change in distance which is a fraction of the width of a proton. 
%changed the relative length of the two arms by $10^{-21}$, a distance smaller than the width of a proton~\cite{GW150914}.

\begin{comment}
Gravitational wave detectors such as LIGO and Virgo are large, complex experiments. However, their design is based fundamentally on a Michelson interferometer, an optical configuration commonly used in undergraduate laboratories.
It consists of a beam splitter which splits light from a laser down two perpendicular arms. % I suggest moving this to later as we are talking about Michelson interferometers generally here. 
%($4\,{\rm km}$ and $3\,{\rm km}$ long at LIGO and Virgo respectively).
Mirrors at the end of each arm reflect the light back towards the beam splitter, where they recombine and produce an interference pattern.
This pattern is dependent on the relative distance travelled by each beam and thus on the relative length of each arm. 
Current generation gravitational wave observatories are able to measure minuscule changes in distance due to passing gravitational waves. 
The first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$~\cite{GW150914}, smaller than the width of a proton.
\end{comment}

% I rephrased to lead with the analogy here and removed 'often' as GW signals are only really listened to for public engagement interest - hannah
Sound is a commonly-used analogy when explaining gravitational wave science. 
Gravitational wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound. For example, gravitational waves can propagate in a vacuum, whereas sound cannot. Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
% rephrased as this what we are doing, rather than what scientists in general are doing
%This analogue allows scientists to use sound vibrations as a simple means to move the components of a demonstration interferometer. 
Detection and analysis techniques used by the gravitational wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
The vibrations move the mirrors, changing the length of the interferometer arms, and therefore the interference pattern.
We use audio signals throughout this work to simulate gravitational wave-like signals.
% and  to demonstrate different data analysis and signal processing techniques used to recover such signals.

@@

The optical configuration of the Michelson interferometer used in this work is shown in Fig.~\ref{fig:ifo_schematic_webcam}.
It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses a green eye-safe laser (top-left) with peak emission at a wavelength of $532\,{\rm nm}$.
The output of the laser is first reflected by a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
After passing through a converging (plano-convex) lens with focal length $125.0\,{\rm mm}$, the laser beam is incident on a beamsplitter, which reflects half of the light towards mirror 1 (M1) and transmits the other half to mirror 2 (M2). 
A $0.5\,{\rm W}$ speaker is fixed to the back of M2 using commercial adhesive putty and serves as a controllable source of vibrations. 
This speaker is one of a pair of commercial, USB-powered speakers, fed by a $3.5\,{\rm mm}$ jack and driven by a PC. 
The other speaker in the pair is kept face-down and as far away from the interferometer as possible to prevent it from acting as a second source.
The beams are reflected by mirrors M1 and M2, located at the end of $\approx 7.5\,{\rm cm}$ and $\approx 10.0\,{\rm cm}$ long arms.
The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging (bi-concave) lens of focal length $-25.0\,{\rm mm}$, and projected onto a screen.
The interference pattern, as shown by the photograph in Fig.~\ref{fig:interference_pattern}, is a set of concentric light and dark rings (fringes). 
A change in the relative length of the arms causes these rings to move radially in or outwards.
Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2020}.
% Ring interference patterns like this are often used in gravitational wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
% Why remove the sentence above? Rings are chosen rather than happening by default. - Hannah
The intensity timeseries is recorded via webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
In Section~\ref{sec:optical_microphone}, which considers more complex audio, the webcam is replaced with a photodiode, which offers a higher sample rate than the webcam. 
%rephrased as this sentence is confusing I think: An intensity time series of some point in the interference pattern is recorded by a webcam (or a photodiode in Section~\ref{sec:optical_microphone}) for data analysis to extract the signal injected at the speaker.

\begin{figure}
 \begin{center}
  \includegraphics[height=0.3\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
 \end{center}
 \caption{\label{fig:interference_pattern}
The interference pattern produced by the table-top Michelson interferometer.
This image was taken with the webcam off the beam axis, which may explain why the (predicted circular) pattern appears elliptical.
The pink star indicates the point where the intensity time series measurements were taken.
The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
}
\end{figure}

>>>>

Gravitational-wave detectors such as LIGO and Virgo are large, complex experiments. 
However, their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
In a Michelson interferometer, laser light is split by a beamsplitter into two perpendicular arms (see Fig.~\ref{fig:ifo_schematic_webcam}). 
Mirrors at the end of each arm reflect the two beams back to the beamsplitter where they recombine to produce an interference pattern.
The resulting interference pattern is dependent on the relative distance travelled by the beams. 
Current generation gravitational-wave observatories have kilometre-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
They can measure minuscule changes in distance due to gravitational waves, e.g. the first detection of a binary black hole merger produced a strain of $10^{-21}$, corresponding to a change in distance which is a fraction of the width of a proton.

Sound is a commonly-used analogy when explaining gravitational-wave science. 
Gravitational-wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound. For example, gravitational waves can propagate in a vacuum, whereas sound cannot. Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
The vibrations move the mirrors, changing the length of the interferometer arms, and therefore the interference pattern.
We use audio signals throughout this work to simulate gravitational-wave-like signals.

@@

The optical configuration of the Michelson interferometer used in this work is shown in Fig.~\ref{fig:ifo_schematic_webcam}.
It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses a green eye-safe laser (top-left) with peak emission at a wavelength of $532\,{\rm nm}$.
The output of the laser is first reflected by a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
After passing through a converging (plano-convex) lens with focal length $125.0\,{\rm mm}$, the laser beam is incident on a beamsplitter, which reflects half of the light towards mirror 1 (M1) and transmits the other half to mirror 2 (M2). 
A $0.5\,{\rm W}$ speaker is fixed to the back of M2 using commercial adhesive putty and serves as a controllable source of vibrations. 
This speaker is one of a pair of commercial, USB-powered speakers, fed by a $3.5\,{\rm mm}$ jack and driven by a PC. 
The other speaker in the pair is kept face-down and as far away from the interferometer as possible to prevent it from acting as a second source.
The beams are reflected by mirrors M1 and M2, located at the end of $\approx 7.5\,{\rm cm}$ and $\approx 10.0\,{\rm cm}$ long arms.
The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging (bi-concave) lens of focal length $-25.0\,{\rm mm}$, and projected onto a screen.
The interference pattern, as shown by the photograph in Fig.~\ref{fig:interference_pattern}, is a set of concentric light and dark rings (fringes). 
A change in the relative length of the arms causes these rings to move radially in or outwards.
Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2020}.
The intensity timeseries is recorded via webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
In Section~\ref{sec:optical_microphone}, which considers more complex audio, the webcam is replaced with a photodiode, which offers a higher sample rate than the webcam.

\begin{figure}
 \begin{center}
  \includegraphics[height=0.3\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
 \end{center}
 \caption{\label{fig:interference_pattern}
The interference pattern produced by the table-top Michelson interferometer.
This image was taken with the webcam off the beam axis.%, which may explain why the (predicted circular) pattern appears elliptical.
The pink star indicates the point where the intensity time series measurements were taken.
The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
}
\end{figure}
